---
title: "Bechdel Project"
output: html_document
date: "2024-11-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# install packages
install.packages("pacman")
pacman::p_load(tidyverse, readr, ggplot2, dplyr, tidyr)
```
## Loading data
```{r}
## Load data
bechdel_df <- read_csv("../data/bechdel_movies_2023_FEB.csv")
imdb_df <- read_csv("../data/IMDb_Dataset_3.csv")

head(bechdel_df)
head(imdb_df)
```
Info:
The bechdel test data is from the bechdeltest.com website API scraped originally. Then it is collated and downloaded from someone else on Kaggle. It only shows the movies that have been approved by the bechdeltest.com website. 

```{r}
bechdel_df$year <- as.numeric(bechdel_df$year)
imdb_df$Year <- as.numeric(imdb_df$Year)

# Plot number of films for each release year, see every 5 years on x-axis for each data set
bechdel_df %>%
  count(year) %>%
  ggplot(aes(x = year, y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Number of films for each release year",
       x = "Release Year",
       y = "Count") +
  scale_x_continuous(breaks = seq(1900, 2020, by = 5))  # Show every 5 years on x-axis

# All column name spaces converted to underscore
colnames(imdb_df) <- gsub(" ", "_", colnames(imdb_df))
imdb_df %>%
  count(Year) %>%
  ggplot(aes(x = Year, y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Number of films for each release year",
       x = "Release Year",
       y = "Count") +
  scale_x_continuous(breaks = seq(1900, 2020, by = 5))  # Show every 5 years on x-axis
```


## Data Cleaning
```{r}
# rename title in imdb
imdb_df <- imdb_df %>% rename(title = Title)
bechdel_df <- bechdel_df %>% rename(Year = year)

# lowercase all titles in bechdel and in imdb df
bechdel_df$title <- tolower(bechdel_df$title)
imdb_df$title <- tolower(imdb_df$title)


# Remove all punctuation from titles
bechdel_df$title <- gsub("[[:punct:]]", "", bechdel_df$title)
imdb_df$title <- gsub("[[:punct:]]", "", imdb_df$title)

## Remove all numbers from titles
bechdel_df$title <- gsub("[0-9]", "", bechdel_df$title)
imdb_df$title <- gsub("[0-9]", "", imdb_df$title)

# Remove any leading spaces from dfs
bechdel_df$title <- str_trim(bechdel_df$title, side = "left")
imdb_df$title <- str_trim(imdb_df$title, side = "left")

# merge all data frames based on title and Year columns
dfs <- list(bechdel_df, imdb_df)
df <- reduce(dfs, inner_join, by = c("title", "Year"))



# Remove following columns: Poster-src, Star_Cast, Overview, Duration_(minutes), Poster_src, date, year, submitterid, id
df <- df %>% 
  subset(select = -c(date, id, Star_Cast, `Duration_(minutes)`, `Poster-src`, submitterid))

head(df)

# All column name spaces converted to underscore
colnames(df) <- gsub(" ", "_", colnames(df))
```

Columns:

IMDB rating - 
The average rating of the movie on IMDB from IMDB users - each user gets one vote. - https://help.imdb.com/article/imdb/track-movies-tv/ratings-faq/G67Y87TFYYP6TWAV#

id - 
The bechdeltest.com unique id.

rating - 
The actual score. Number from 0 to 3 (0 means no two women, 1 means no talking, 2 means talking about a man, 3 means it passes the test)

dubious - 
If the submitter considers the score is dubious on the bechdeltest.com

imdbid - 
The ID of the movie on IMDB

visible - 
Has this movie been approved (currently only approved movies are returned, so this value will always be 1)

Genre -
The genre of the movie - IMDB users are able to edit and add these. Multiple genres are sometimes added

Certificate - 
Meta_score - 
No_of_Votes - 
Gross - 



## Preliminary plotting of data
```{r}
# Make table with no. of films with each rating value of 0-3
df %>% 
  group_by(rating) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = rating, y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Number of films with each rating value",
       x = "Rating",
       y = "Count")
```

```{r}
# plot number of films in each genre category across columns "genre1" "genre2" "genre3"
df %>%
  gather(key = "genre_column", value = "genre_value", Genre:Third_Genre) %>%  # Gather all genre columns
  filter(!is.na(genre_value)) %>%  # Remove rows where genre is NA
  count(genre_value, name = "Count") %>%  # Count occurrences of each genre
  ggplot(aes(x = reorder(genre_value, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Number of films in each genre category",
       x = "Genre",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


## Facetwrap by Bechdel rating
df %>%
  gather(key = "genre_column", value = "genre_value", Genre:Third_Genre) %>%  # Gather all genre columns
  filter(!is.na(genre_value)) %>%  # Remove rows with NA genres
  count(rating, genre_value, name = "Count") %>%  # Count occurrences of each genre per rating
  ggplot(aes(x = reorder(genre_value, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Number of films in each genre category by Rating",
       x = "Genre",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~rating,ncol=1)  # Facet by Rating
```


```{r}
# Remove all non-numeric characters in release year values
df$Year <- gsub("[^0-9]", "", df$Year)
# Look at all unique values in release year and see in numerical order
df$Year <- as.numeric(df$Year)
unique(df$Year)


# Plot number of films for each release year, see every 5 years on x-axis
df %>%
  count(Year) %>%
  ggplot(aes(x = Year, y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Number of films for each release year",
       x = "Release Year",
       y = "Count") +
  scale_x_continuous(breaks = seq(1900, 2020, by = 5))  # Show every 5 years on x-axis
```
